---
title: "US Storm Consequences"
author: "Phillip Chaffee"
date: "Wednesday, July 22, 2015"
output: html_document
---

##Synopsis

This analysis attempts to use the National Weather Service's Storm Data Documentation to answer two questions about US storms.

Which storm types are the most harmful with respect to population health?

Which storm types have the greatest economic consequences?

The study found that the storm types with the greatest health consequences are tornado, thunderstorm wind, flood, flash flood, heat, excessive heat, and lightning.

The storm types with the greatest economic consequences are drought, flood, flash flood, hail, extreme cold, frost/freeze, tornado, and hurricane.


##Data Processing

In this section we will take our raw data and transform it into something we can easily work with to answer our questions.

First, we will download and store the data.

```{r download and store, cache=TRUE}
download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2", destfile="StormData.csv.bz2")

stormdata <- read.csv(bzfile("StormData.csv.bz2"), stringsAsFactors=F)
```

What variables do we have to work with?

```{r storm variables, cache=TRUE}
colnames(stormdata)
head(stormdata)
```

We need the variables that have to do with storm type, health effects, and economic effects. These seem to be EVTYPE, FATALITIES, INJURIES, PROPDMG, PROPDMGEXP, CROPDMG, and CROPDMGEXP. So let's subset to a data set with only those variables.

```{r subset dataset, cache=TRUE}
columns <- c("EVTYPE", "FATALITIES", "INJURIES", "PROPDMG", "PROPDMGEXP", "CROPDMG", "CROPDMGEXP")
stormdata <- stormdata[,columns]
head(stormdata)
```

It also seems that the damage numbers need to be multiplied by their corresponding exponent. Let's do that and shrink the data set even more.

```{r damage variables, cache=TRUE}
for(i in 1:length(stormdata$EVTYPE)) {
  if(stormdata[i,5]=="K"){
    stormdata[i,4] <- stormdata[i,4]*1000
  }else if(stormdata[i,7]=="K"){
    stormdata[i,6] <- stormdata[i,6]*1000
  }
  if(stormdata[i,5]=="M"){
    stormdata[i,4] <- stormdata[i,4]*1000000
  }else if(stormdata[i,7]=="M"){
    stormdata[i,6] <- stormdata[i,6]*1000000
  }
}

columns <- c("EVTYPE", "FATALITIES", "INJURIES", "PROPDMG", "CROPDMG")
stormdata <- stormdata[,columns]
head(stormdata)
```

We now have a clean data set with only the variables we need to work with.

##Data Analysis

Which storm types are the most harmful with respect to population health?

```{r population health, cache=TRUE}
injuries <- aggregate(INJURIES ~ EVTYPE, sum, data=stormdata, na.action=na.omit)
fatalities <- aggregate(FATALITIES ~ EVTYPE, sum, data=stormdata, na.action=na.omit)

injuries<- injuries[order(-injuries$INJURIES),]
injuries<- injuries[1:5,]

fatalities <- fatalities[order(-fatalities$FATALITIES),]
fatalities <- fatalities[1:5,]
```

Which storm types have the greatest economic consequences?

```{r economic health, cache=TRUE}
propdmg <- aggregate(PROPDMG ~ EVTYPE, sum, data=stormdata, na.action=na.omit)
cropdmg <- aggregate(CROPDMG ~ EVTYPE, sum, data=stormdata, na.action=na.omit)

propdmg <- propdmg[order(-propdmg$PROPDMG),]
propdmg <- propdmg[1:5,]

cropdmg <- cropdmg[order(-cropdmg$CROPDMG),]
cropdmg <- cropdmg[1:5,]
```

##Results

####Which storm types are the most harmful with respect to population health?

```{r harm graphs}
library(lattice)
barchart(INJURIES ~ EVTYPE, injuries)
```

The top five storms types for injuries are tornado, thunderstorm wind, lightning, flood, and excessive heat.

The fatality data is a bit different, but mostly the same. Here are the exact numbers.

```{r harm results}
injuries
fatalities
```



####Which storm types have the greatest economic consequences?

```{r economic graphs}
barchart(CROPDMG ~ EVTYPE, cropdmg)
```
The worst thing for crops is drought.

Let's take a look at the exact numbers on crop damage and property damage by storm type.

```{r economic results}
cropdmg
propdmg
```
