---
title: "ToothGrowth Analysis"
author: "Phillip Chaffee"
date: "Sunday, July 26, 2015"
output: pdf_document
---

#Synopsis

This report takes a look at the dataset within R on tooth growth. What it is, and some comparisons of the different vairables effects on tooth growth.

#Data Analysis

####Load the data

```{r load}
library(datasets)
toothgrowth <- ToothGrowth
```

####Analyze and summarize the data

```{r data}
colnames(toothgrowth)
head(toothgrowth)
```

The data set measures tooth growth in length (len) as affected by Vitamin C administration (supp), given through ascorbic acid (VC) and orange juice (OJ) in variant doses (dose).

####Explore the data

Let's look at how tooth growth relates to Vitamin C dose between type of administration: orange juice or asorbuic acid.

```{r admin}
library(lattice)
xyplot(len ~ dose | supp, data=toothgrowth, type=c("p","r"))
```

##Conclusion

It seems that Vitamin C affects tooth growth no matter what, but does a better job in small doses when administered through Orange Juice.